An Embedding Framework for the Design and Analysis of Consistent Polyhedral Surrogates

### 2022-06-28 poly-embed-journal/poly-embed-journal.tex
meeting notes
getting close on E

\begin{lemma}\label{lem:G-unique-facets-Hstar}
  %Given a finite set $\V \subset \reals^\Y_+$, define $\H = \H_\V \cup \H_\Y$.
  %\raf{This is the important general polyhedra result to cite/state.  Note: we'll have to work half-spaces of the form $H_{(w,\beta)}^+ = \{ z \in \reals^{n+1} \mid \inprod{z}{w} - \beta \geq 0\}$.}  
  Consider a loss $L : \reals^d \to \reals^\Y_+$ with polyhedral extended risk $\risk L_+$.
  There is a unique irredundant set $\H^*$ such that $\hyp(g_{L(\R)}) = \cap \H^*$. 
  Moreover, for each $H^+ \in\H^*$ and $H$ such that $H = \partial H^+$, the face $\hyp(g_{L(\R)}) \cap H$ is a facet. 
  Moreover, $\H_\Y \subseteq \H^*$. 
\end{lemma}
\begin{proof}
	First, we need to show that $\hyp(g_{L(\R)})$ is full-dimensional.
	Since $g_{L(\R)}$ is nonnegative on $\reals_+^\Y$, $\hyp(g_{L(\R)})$ therefore contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is $(|\Y| + 1)$-dimensional.

  Since $\hyp(g_{L(\R)})$ is full-dimensional, take $\H^*$ to be the unique irredundant and finite set of closed halfspaces such that $\hyp(g_{L(\R)}) = \cap \H^*$ from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(i).
  Now, $\hyp(g_{L(\R)}) \cap H$ for any $H \in \H^*$ being a facet follows immediately from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(ii) and (iii).

  To show that $\H_\Y \subseteq \H^*$, it suffices from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(ii) to show that each $F_y := H_y \cap \hyp(g_{L(\R)})$ is a facet.
  First, observe that $F_y$ is a face of $\hyp(g_{L(\R)})$ as $H_y$ supports $\hyp(g_{L(\R)})$ at the origin.\jessiet{Need to justify?}
  Moreover, $F_y = H_{(e_y, 0)} \cap \hyp(g_{L(\R)}) = \{(x,c) \mid x_y = 0\} \cap \hyp(g_{L(\R)})$ is a $|\Y|$-dimensional face as it contains $\{(x,c) \mid x_y = 0, c \leq 0\}$, which is $|\Y|$-dimensional.
%  To see $\H_\Y \subseteq \H^*$, if there was a $y \in \Y$ such that $H^+_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, it is either because it is redundant or $H^+_y \not \subseteq \hyp(g_{L(\R)})$.
%  In the first case, there must be some $\epsilon > 0$ such that $\{(x,c) \mid x_y \geq \epsilon\} \in \H^*$, which cannot happen as $g_{L(\R)}$ is concave and finite (namely, nonnegative) at $x = e_y$.
%  In the second case, we have a point $x \not \in \reals^\Y_+$ with $g_{L(\R)}(x) > -\infty$.
%  This contradicts the fact that $g_{L(\R)}$ is only finite on $\reals^\Y_+$ by construction with the convex indicator. \raft{TODO: Couldn't follow this last line, since $\delta$ can be $\infty$\jessie{does it help that it's minus the convex indicator?} 6/20/2022: still don't get it, sorry :-)... \jessie{Yeah I changed the sentence entirely.  better? \# burn}}. 
\end{proof}

@@

\begin{claim}\label{claim:extended-poly-risk-poly}
	For any minimizable function $L$ such that $\risk L$ is polyhedral, its extension $\risk L_+$ is also polyhedral.
\end{claim}
\begin{proof}
  We will think of $\risk L$ as defined $\risk L:\reals^\Y\to\reals_+\cup\{-\infty\}$  with $\dom(\risk L) = \simplex$.
For $p \in \simplex$, we know $\sum_i p_i = 1$, and can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
If $p \not \in \simplex$, then $\risk L(p) = -\infty$ and inner products are not used to compute $\risk L$.
Moreover, since $\risk L$ is polyhedral, it is finitely generated \citep[Proposition 19.1.2]{rockafellar1997convex} and can be written 
\begin{align*}
\risk L(p) 
&= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex)\\
&= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
\end{align*}
Therefore, it is positively $1$-homogeneous, and its extension $\risk L_+$ is polyhedral immediately.
\end{proof}

@@

\lemmaX*
\begin{proof}
	%\jessiet{Get to Assumption~\ref{assum:V-star-exists-infinite} quickly}
%As in this appendix, we consider $\V = L(\R)$ such that $g_\V = \risk L_+$.
%\jessie{$\risk L$ to $\risk L_+$.  Change Claim 6: if you have a minimizable loss and $\risk L$ is polyhedral, then $\risk L_+$ is polyhedral.}
Since $\risk L$ is polyhedral, so is $\risk L_+$ by Claim~\ref{claim:extended-poly-risk-poly}.
Therefore, we have satisfied the requirements of Proposition~\ref{prop:poly-risk-assumption-satisfied}, and can conclude there is a $\V$ such that $\risk L_+ = g_{\V}$ satisfying Assumption~\ref{assum:V-star-exists-infinite}.
Moreover, there must be a finite set $\R^*$ such that $\V = L(\R^*)$ by construction, which is representative as some element of $\R^*$ generates the Bayes risk for every $p \in \simplex$ as $\risk L = \min_{v \in L(\R^*)}\inprod{p}{v} - \delta\{p \mid \simplex\}$.
%Since $\risk L$ is polyhedral and concave there exists a finite set $\V$ such that $\risk L = f_\V$ by Claim~\ref{claim:f-min-affine}.
%Moreover as $\V$ is finite, $g_{\V} = \risk L_+$ (by $1$-homogeneity following from the form of $g_{\V}$) is also polyhedral and concave, and $f_\V = g_\V$ on $\simplex$.
%Therefore, there is a $\V^* \subseteq \V$ such that $L$ and $\V^*$ satisfy Assumption~\ref{assum:V-star-exists-infinite} by Proposition~\ref{prop:poly-risk-assumption-satisfied}.
%Moreover, any set $\R$ such that $\V^* = L(\R)$ is a finite representative set for $L$, and such a set must exist by the existence and construction of $\V^*$.
%%Given the form of $\risk L$ written as a pointwise max of finitely many affine functions (which are loss vectors) in $\V$, the loss $L$ has a finite representative set. 
%%Moreover, by Claim~\ref{claim:gV-and-fV}, $g_\V = f_\V$ on the simplex.

Now that we have the preliminaries of $L, \V$ satisfying Assumption~\ref{assum:V-star-exists-infinite}, consider the itemized statements.
Lemma~\ref{lem:lemX1-rep-iff-subset-vectors} is exactly statement \eqref{item:X-rep-V}.
This immediately implies statement~\eqref{item:X-min-V}.
Moreover, Lemma~\ref{lem:lemX-3-rep-iff-FDLS-subsets} is exactly statement~\eqref{item:X-rep-Theta}, and again statement~\eqref{item:X-min-Theta} immediately follows.
Statement~\eqref{item:X-rep-contain-min} is a corollary of the existence of a finite representative set, which follows since $\V \subseteq L(\R)$.
Statement~\eqref{item:X-full-dim} is exactly Lemma~\ref{lem:fdls-exactly-theta}.
Statement~\eqref{item:X-redundant} is exactly Lemma~\ref{lem:any-levelset-contained-in-minlevelset}.
Finally, Statement~\eqref{item:X-tight-embed} follows as a corollary of statement~\eqref{item:X-min-V} and Corollary~\ref{cor:tight-embed-min-rep}.
\end{proof}

>>>>

\begin{lemma}\label{lem:G-unique-facets-Hstar}
  %Given a finite set $\V \subset \reals^\Y_+$, define $\H = \H_\V \cup \H_\Y$.
  %\raf{This is the important general polyhedra result to cite/state.  Note: we'll have to work half-spaces of the form $H_{(w,\beta)}^+ = \{ z \in \reals^{n+1} \mid \inprod{z}{w} - \beta \geq 0\}$.}  
  Consider a loss $L : \reals^d \to \reals^\Y_+$ with polyhedral extended risk $\risk L_+$.
  There is a unique irredundant set $\H^*$ of closed halfspaces such that $\hyp(g_{L(\R)}) = \cap \H^*$. 
  Moreover, for each $H^+ \in\H^*$ and $H$ such that $H = \partial H^+$, the face $\hyp(g_{L(\R)}) \cap H$ is a facet. 
  Moreover, $\H_\Y \subseteq \H^*$. 
\end{lemma}
\begin{proof}
	As $g_{L(\R)}$ is nonnegative on $\reals_+^\Y$, the set $\hyp(g_{L(\R)})$ therefore contains $\{(x,c) \mid x\in\reals^\Y_+, c \leq 0\}$, which is $(|\Y| + 1)$-dimensional.
  Therefore $\hyp(g_{L(\R)})$ is full-dimensional.
	Take $\H^*$ to be the unique irredundant and finite set of closed halfspaces such that $\hyp(g_{L(\R)}) = \cap \H^*$ from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(i).
  Now, $\hyp(g_{L(\R)}) \cap H$ for any $H \in \H^*$ being a facet follows immediately from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(ii) and (iii).

  To show that $\H_\Y \subseteq \H^*$, it suffices from Theorem~\ref{thm:polyhedron-uniquely-gen-facets}(ii) to show that each $F_y := H_y \cap \hyp(g_{L(\R)})$ for $y\in\Y$ is a facet.
  From Claim~\ref{claim:gV-equals-riskL}, since $H_y^+ \in \H_\Y$, we have $\hyp(g_{L(\R)}) \subseteq H_y^+$.
  We have $F_y = H_{(e_y, 0)} \cap \hyp(g_{L(\R)}) = \{(x,c) \mid x_y = 0, c \leq g_{L(\R)}(x)\} = \{(x,c) \mid x\in\reals^\Y_+, x_y=0, c \leq g_{L(\R)}(x)\} \supseteq \{(x,c) \mid x\in\reals^\Y_+, x_y=0, c \leq 0\}$ as $g_{L(\R)}(x) \geq 0$.
  Thus, $F_y$ is a nonempty face of $\hyp(g_{L(\R)})$, and contains a $|\Y|$-dimensional set, so must be a facet.
%  To see $\H_\Y \subseteq \H^*$, if there was a $y \in \Y$ such that $H^+_y = \{(x,c) \mid x_y \geq 0\}$ was not in $\H^*$, it is either because it is redundant or $H^+_y \not \subseteq \hyp(g_{L(\R)})$.
%  In the first case, there must be some $\epsilon > 0$ such that $\{(x,c) \mid x_y \geq \epsilon\} \in \H^*$, which cannot happen as $g_{L(\R)}$ is concave and finite (namely, nonnegative) at $x = e_y$.
%  In the second case, we have a point $x \not \in \reals^\Y_+$ with $g_{L(\R)}(x) > -\infty$.
%  This contradicts the fact that $g_{L(\R)}$ is only finite on $\reals^\Y_+$ by construction with the convex indicator. \raft{TODO: Couldn't follow this last line, since $\delta$ can be $\infty$\jessie{does it help that it's minus the convex indicator?} 6/20/2022: still don't get it, sorry :-)... \jessie{Yeah I changed the sentence entirely.  better? \# burn}}. 
\end{proof}

@@

\begin{claim}\label{claim:extended-poly-risk-poly}
	For any minimizable function $L$ such that $\risk L$ is polyhedral, its extension $\risk L_+$ is also polyhedral.
\end{claim}
\begin{proof}
  We will think of $\risk L$ as defined $\risk L:\reals^\Y\to\reals_+\cup\{-\infty\}$  with $\dom(\risk L) = \simplex$.
For $p \in \simplex$, we know $\sum_i p_i = 1$, and can write any inner product $\inprod{p}{b} - \beta = \inprod{p}{b} - \inprod{p}{\beta \ones} = \inprod{p}{b - \beta \ones}$.
If $p \not \in \simplex$, then $\risk L(p) = -\infty$.
Moreover, since $\risk L$ is polyhedral, it is finitely generated \citep[Proposition 19.1.2]{rockafellar1997convex} and can be written 
\begin{align*}
\risk L(p) 
&= \min(\inprod{p}{b_1} - \beta_1, \ldots, \inprod{p}{b_k} - \beta_k) - \delta(p \mid \simplex)\\
&= \min(\inprod{p}{b_1 - \beta_1 \ones}, \ldots, \inprod{p}{b_k - \beta_k \ones}) - \delta(p \mid \simplex)~.~
\end{align*}
Therefore, it is positively $1$-homogeneous, and its extension $\risk L_+$ is polyhedral immediately\raft{Is the ``it'' $\risk L$?  Since then it's not $1$-homog, its extension is.  You have to connect the display to the conclusion; one way is to write out what $\risk L_+$ is, but you have to justify it either way.}.
\end{proof}

@@

\lemmaX*
\begin{proof}
	%\jessiet{Get to Assumption~\ref{assum:V-star-exists-infinite} quickly}
%As in this appendix, we consider $\V = L(\R)$ such that $g_\V = \risk L_+$.
%\jessie{$\risk L$ to $\risk L_+$.  Change Claim 6: if you have a minimizable loss and $\risk L$ is polyhedral, then $\risk L_+$ is polyhedral.}
Since $\risk L$ is polyhedral, so is $\risk L_+$ by Claim~\ref{claim:extended-poly-risk-poly}.
Therefore, we have satisfied the requirements of Proposition~\ref{prop:poly-risk-assumption-satisfied}, and can conclude there is a $\V^*$ such that $\risk L_+ = g_{\V^*}$ satisfying Assumption~\ref{assum:V-star-exists-infinite}.
Moreover, there must be a finite set $\R^*$ such that $\V^* = L(\R^*)$ by construction, which is representative as some element of $\R^*$ generates the Bayes risk for every $p \in \simplex$ as $\risk L = \min_{v \in L(\R^*)}\inprod{p}{v} - \delta\{p \mid \simplex\}$\raft{``generates the Bayes risk'' not defined; need to slow down here I think and find the rigorous statement among the many things you proved!}.
%Since $\risk L$ is polyhedral and concave there exists a finite set $\V$ such that $\risk L = f_\V$ by Claim~\ref{claim:f-min-affine}.
%Moreover as $\V$ is finite, $g_{\V} = \risk L_+$ (by $1$-homogeneity following from the form of $g_{\V}$) is also polyhedral and concave, and $f_\V = g_\V$ on $\simplex$.
%Therefore, there is a $\V^* \subseteq \V$ such that $L$ and $\V^*$ satisfy Assumption~\ref{assum:V-star-exists-infinite} by Proposition~\ref{prop:poly-risk-assumption-satisfied}.
%Moreover, any set $\R$ such that $\V^* = L(\R)$ is a finite representative set for $L$, and such a set must exist by the existence and construction of $\V^*$.
%%Given the form of $\risk L$ written as a pointwise max of finitely many affine functions (which are loss vectors) in $\V$, the loss $L$ has a finite representative set. 
%%Moreover, by Claim~\ref{claim:gV-and-fV}, $g_\V = f_\V$ on the simplex.

Now consider the itemized statements.
Lemma~\ref{lem:lemX1-rep-iff-subset-vectors} is exactly statement \eqref{item:X-rep-V}.
This immediately implies statement~\eqref{item:X-min-V}.
Moreover, Lemma~\ref{lem:lemX-3-rep-iff-FDLS-subsets} is exactly statement~\eqref{item:X-rep-Theta}, and again statement~\eqref{item:X-min-Theta} immediately follows.
Statement~\eqref{item:X-rep-contain-min} is a corollary of the existence of a finite representative set, which follows since $\V^* \subseteq L(\R)$.
Statement~\eqref{item:X-full-dim} is exactly Lemma~\ref{lem:fdls-exactly-theta}.
Statement~\eqref{item:X-redundant} is exactly Lemma~\ref{lem:any-levelset-contained-in-minlevelset}.
Finally, Statement~\eqref{item:X-tight-embed} follows as a corollary of statement~\eqref{item:X-min-V} and Corollary~\ref{cor:tight-embed-min-rep}.
\end{proof}
